---
title: "AT4BR_Tidy"
format: html
editor: visual
---

## Assignment

Use **tidyr** and **dplyr** libraries whenever possible, to do the following:

1- Load the "iris" dataset.

Dataset is included in base R.

2- Transform it to longer (tip: you should get 3 columns Species \| Trait \| Size )

```{r}
library(tidyr)
library(dplyr)

# Transform iris to longer format
iris_long <- iris %>%
  pivot_longer(
    cols = -Species,
    names_to = "Trait",
    values_to = "Size"
  )
iris_long
```

3- Answer the following questions. (tip: you can answer each question "piping" only 2-3 commands )

A\) What is the **mean** length and width of the sepal and petal of each species?

```{r}
iris_long %>%
  group_by(Species, Trait) %>%
  summarize(
    Mean_Size = mean(Size),
    .groups = 'drop'
  )
```

B\) What is the **largest** and **widest**,  **petal** and **sepal**,  of **virginica** and **versicolor** species?

```{r}
iris_long %>%
  filter(Species %in% c("virginica", "versicolor")) %>%
  group_by(Species, Trait) %>%
  summarize(
    Max_Size = max(Size),
    .groups = 'drop'
  )
```

C\) For each species, for **sepal** and **petal, width** and **length**,  how many observations are above the mean?\*    

```{r}
iris_long %>%
  group_by(Species, Trait) %>%
  mutate(
    Mean_Size = mean(Size)
  ) %>%
  ungroup() %>%
  filter(Size > Mean_Size) %>%
  count(Species, Trait)
```
